---
title: "Sam - Part 1"
author: "Samuel Miller"
date: "12/6/2020"
output: html_document
---

Import the necessary libraries and datasets and filter them down to only contain the information we need

```{r, message=FALSE}
library(dplyr)
library(tidyverse)
library(ggplot2)
library(gganimate)

pop <- read.csv("https://github.com/CSSEGISandData/COVID-19/raw/master/csse_covid_19_data/UID_ISO_FIPS_LookUp_Table.csv") %>% select(c(6:8, 12))
cases <- read.csv("https://github.com/CSSEGISandData/COVID-19/raw/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_US.csv") %>% select(c(6:8, 12:327))
income <- read.csv("income.csv")

us_pop <- pop[pop$Country_Region == "US",]
us_cases <- cases[cases$Country_Region == "US",]

us_cases$Admin2 <- paste(us_cases$Admin2, us_cases$Province_State)
us_pop$Admin2 <- paste(us_pop$Admin2, us_pop$Province_State)
income$County <- paste(income$County, income$State)

us_pop <- us_pop %>% select(c(1, 4))
us_cases <- us_cases %>% select(c(1, 4:319))
income <- income %>% select(c(2:3))
```

Separates the income dataset into income brackets, storing the data in buckets

```{r, message=FALSE}
separate <- function(x) {
  if(x < 36000) 1
  else if(x < 41000) 2
  else if(x < 46500) 3
  else if(x < 54000) 4
  else 5
}

max_val <- max(income$Income)
buckets <- vector(mode = "list", length = 5)

for(i in 1:length(income$Income)) {
  idx <- separate(income$Income[[i]])
  buckets[[idx]] <- c(buckets[[idx]], income$County[[i]])
}
```

Synthesizes and transforms the data so that we have a dataset that is best formatted for graphing the information we want to (date, cumulative cases per capita, and income bracket), then graphs it

```{r, message=FALSE}
us_cases_per_capita <- data.frame(t((us_cases %>% select(-c(316, 317)))[-1]))
colnames(us_cases_per_capita) <- us_cases[, 1]
us_cases_per_capita <- us_cases_per_capita[, names(us_cases_per_capita) %in% income$County]

b <- list(rep(0, 314), rep(0, 314), rep(0, 314), rep(0, 314), rep(0, 314))
l <- list(length(buckets[[1]]), length(buckets[[2]]), length(buckets[[3]]), length(buckets[[4]]), length(buckets[[5]]))

for(county in names(us_cases_per_capita))
  for(i in 1:5)
    if(county %in% buckets[[i]])
      b[[i]] <- b[[i]] + us_cases_per_capita[[county]] / us_pop_filtered[[county]]

dates <- rep(1:314, 5)

all_cases <- c(b[[1]] / l[[1]], b[[2]] / l[[2]], b[[3]] / l[[3]], b[[4]] / l[[4]], b[[5]] / l[[5]])
incomes <- c(rep(36000, 314), rep(41000, 314), rep(46500, 314), rep(54000, 314), rep(54001, 314))

bucketed_data <- data.frame(dates, all_cases, incomes)

p <- ggplot(bucketed_data, aes(dates, all_cases, group = incomes, color = factor(incomes))) +
  geom_line() +
  scale_color_viridis_d() +
  labs(x = "Date", y = "% of the Population Infected", color = "Median Household Income") +
  theme(legend.position = "top") + 
  scale_x_continuous(breaks = seq(7, 300, 29.3), labels=c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov")) +
  scale_color_discrete(breaks = c(36000, 41000, 46500, 54000, 54001), labels=c("< $36,000", "< $41,000", "< $46,500", "< $54,000", "> $54,000")) +
  ggtitle("Average Cumulative Cases") + transition_reveal(dates)

animate(p, duration = 10, end_pause = 75)
```

Synthesizes and transforms the data so that we have a dataset that is best formatted for graphing the information we want to (date, 4-day case totals per capita, and income bracket), then graphs it

```{r, message=FALSE}
us_cases_per_capita <- data.frame(t(us_cases[-1]))
colnames(us_cases_per_capita) <- us_cases[, 1]
us_cases_per_capita <- us_cases_per_capita[, names(us_cases_per_capita) %in% income$County]

us_pop_filtered <- data.frame(t(us_pop[-1]))
colnames(us_pop_filtered) <- us_pop[, 1]
us_pop_filtered <- us_pop_filtered[, names(us_pop_filtered) %in% income$County]

b <- list(rep(0, 79), rep(0, 79), rep(0, 79), rep(0, 79), rep(0, 79))
l <- list(length(buckets[[1]]), length(buckets[[2]]), length(buckets[[3]]), length(buckets[[4]]), length(buckets[[5]]))

for(county in names(us_cases_per_capita))
  for(i in 1:5)
    if(county %in% buckets[[i]])
      b[[i]] <- b[[i]] +  colSums(matrix(c(us_cases_per_capita[[county]][1], diff(us_cases_per_capita[[county]])) / us_pop_filtered[[county]], nrow = 5))

dates <- rep(1:79, 5)
all_cases <- c(b[[1]] / l[[1]], b[[2]] / l[[2]], b[[3]] / l[[3]], b[[4]] / l[[4]], b[[5]] / l[[5]])
incomes <- c(rep(36000, 79), rep(41000, 79), rep(46500, 79), rep(54000, 79), rep(54001, 79))

bucketed_data <- data.frame(dates, all_cases, incomes)

p2 <- ggplot(bucketed_data, aes(dates, all_cases, group = incomes, color = factor(incomes))) +
  geom_line() +
  scale_color_viridis_d() +
  labs(x = "Month", y = "% of the Population Infected", color = "Median Household Income") +
  theme(legend.position = "top") + 
  scale_x_continuous(breaks = seq(7, 300, 29.3), labels=c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov")) +
  scale_color_discrete(breaks = c(36000, 41000, 46500, 54000, 54001), labels=c("< $36,000", "< $41,000", "< $46,500", "< $54,000", "> $54,000")) +
  ggtitle("Average Cases Every 4 Days") + 
  geom_point(aes(group = seq_along(dates))) + 
  transition_reveal(dates)

animate(p2, duration = 10, end_pause = 75)
```
