---
title: "Sam - Part 1"
author: "Samuel Miller"
date: "12/6/2020"
output: html_document
---

Import the necessary libraries and datasets and filter them down to only contain the information we need

```{r, message=FALSE}
library(dplyr)
library(tidyverse)
library(ggplot2)
library(gganimate)

pop <- read.csv("https://github.com/CSSEGISandData/COVID-19/raw/master/csse_covid_19_data/UID_ISO_FIPS_LookUp_Table.csv") %>% select(c(6:8, 12))
cases <- read.csv("https://github.com/CSSEGISandData/COVID-19/raw/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_US.csv") %>% select(c(6:8, 12:327))
income <- read.csv("income.csv")

us_pop <- pop[pop$Country_Region == "US",]
us_cases <- cases[cases$Country_Region == "US",]

us_cases$Admin2 <- paste(us_cases$Admin2, us_cases$Province_State)
us_pop$Admin2 <- paste(us_pop$Admin2, us_pop$Province_State)
income$County <- paste(income$County, income$State)

us_pop <- us_pop %>% select(c(1, 4))
us_cases <- us_cases %>% select(c(1, 4:319))
income <- income %>% select(c(2:3))
```

Separates the income dataset into income brackets, storing the data in buckets

```{r, message=FALSE}
separate <- function(x) {
  if(x < 36000) 1
  else if(x < 41000) 2
  else if(x < 46500) 3
  else if(x < 54000) 4
  else 5
}

max_val <- max(income$Income)
buckets <- vector(mode = "list", length = 5)

for(i in 1:length(income$Income)) {
  idx <- separate(income$Income[[i]])
  buckets[[idx]] <- c(buckets[[idx]], income$County[[i]])
}
```

Synthesizes and transforms the data so that we have a dataset that is best formatted for graphing the information we want to (date, cumulative cases per capita, and income bracket), then graphs it

```{r, message=FALSE}
us_cases_per_capita <- data.frame(t((us_cases %>% select(-c(316, 317)))[-1]))
colnames(us_cases_per_capita) <- us_cases[, 1]
us_cases_per_capita <- us_cases_per_capita[, names(us_cases_per_capita) %in% income$County]

b <- list(rep(0, 314), rep(0, 314), rep(0, 314), rep(0, 314), rep(0, 314))
l <- list(length(buckets[[1]]), length(buckets[[2]]), length(buckets[[3]]), length(buckets[[4]]), length(buckets[[5]]))

for(county in names(us_cases_per_capita))
  for(i in 1:5)
    if(county %in% buckets[[i]])
      b[[i]] <- b[[i]] + us_cases_per_capita[[county]] / us_pop_filtered[[county]]

dates <- rep(1:314, 5)

all_cases <- c(b[[1]] / l[[1]], b[[2]] / l[[2]], b[[3]] / l[[3]], b[[4]] / l[[4]], b[[5]] / l[[5]])
incomes <- c(rep(36000, 314), rep(41000, 314), rep(46500, 314), rep(54000, 314), rep(54001, 314))

bucketed_data <- data.frame(dates, all_cases, incomes)

p <- ggplot(bucketed_data, aes(dates, all_cases, group = incomes, color = factor(incomes))) +
  geom_line() +
  scale_color_viridis_d() +
  labs(x = "Date", y = "% of the Population Infected", color = "Median Household Income") +
  theme(legend.position = "top") + 
  scale_x_continuous(breaks = seq(7, 300, 29.3), labels=c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov")) +
  scale_color_discrete(breaks = c(36000, 41000, 46500, 54000, 54001), labels=c("< $36,000", "< $41,000", "< $46,500", "< $54,000", "> $54,000")) +
  ggtitle("Average Cumulative Cases") + transition_reveal(dates)

animate(p, duration = 10, end_pause = 75)
```

Calculates a line of best fit to model the average cumulative cases for the different income groups. 
To do this, it uses the lm(), I(), and predict() functions.
Note: this function that it calculates models the data well, but realistically, the line should be logarithmic since the data will eventually plateau once enough people get the virus or a vaccine is created.


```{r}
library(reshape2)

########## INCOME GROUP 1 ##########
exp_reg_1 <- lm(y_36000 ~ x_36000 + I(x_36000^2) + I(x_36000^3), data = xy_36000)

prd <- data.frame(x_36000 = seq(0, 350, by = 0.5))

result <- prd
result$exp_reg_1 <- predict(exp_reg_1, newdata = prd)

result <- melt(result, id.vars = "x_36000", variable.name = "model_1",
                value.name = "cases_1")

ggplot(result, aes(x = x_36000, y = cases_1)) +
  theme_bw() +
  geom_point(data = xy_36000, aes(x = x_36000, y = y_36000), size = 0.1) +
  geom_line(aes(colour = model_1), size = 1, color = "red") +
  labs(x = "Day", y = "% of the Population Infected") +
  ggtitle("Average Cumulative Cases \nfor Median Household Income <$36,000")


########## INCOME GROUP 2 ##########
exp_reg_2 <- lm(y_41000 ~ x_41000 + I(x_41000^2) + I(x_41000^3), data = xy_41000)

prd <- data.frame(x_41000 = seq(0, 350, by = 0.5))

result <- prd
result$exp_reg_2 <- predict(exp_reg_2, newdata = prd)

result <- melt(result, id.vars = "x_41000", variable.name = "model_2",
                value.name = "cases_2")
ggplot(result, aes(x = x_41000, y = cases_2)) +
  theme_bw() +
  geom_point(data = xy_41000, aes(x = x_41000, y = y_41000), size = 0.1) +
  geom_line(aes(colour = model_2), size = 1, color = "orange") +
  labs(x = "Day", y = "% of the Population Infected") +
  ggtitle("Average Cumulative Cases \nfor Median Household Income $36,000 - $41,000")


########## INCOME GROUP 3 ##########
exp_reg_3 <- lm(y_46500 ~ x_46500 + I(x_46500^2) + I(x_46500^3), data = xy_46500)

prd <- data.frame(x_46500 = seq(0, 350, by = 0.5))

result <- prd
result$exp_reg_3 <- predict(exp_reg_3, newdata = prd)

result <- melt(result, id.vars = "x_46500", variable.name = "model_3",
                value.name = "cases_3")
ggplot(result, aes(x = x_46500, y = cases_3)) +
  theme_bw() +
  geom_point(data = xy_46500, aes(x = x_46500, y = y_46500), size = 0.1) +
  geom_line(aes(colour = model_3), size = 1, color = "green") +
  labs(x = "Day", y = "% of the Population Infected") +
  ggtitle("Average Cumulative Cases \nfor Median Household Income $41,000 - $46,500")


########## INCOME GROUP 4 ##########
exp_reg_4 <- lm(y_54000 ~ x_54000 + I(x_54000^2) + I(x_54000^3), data = xy_54000)

prd <- data.frame(x_54000 = seq(0, 350, by = 0.5))

result <- prd
result$exp_reg_4 <- predict(exp_reg_4, newdata = prd)

result <- melt(result, id.vars = "x_54000", variable.name = "model_4",
                value.name = "cases_4")
ggplot(result, aes(x = x_54000, y = cases_4)) +
  theme_bw() +
  geom_point(data = xy_54000, aes(x = x_54000, y = y_54000), size = 0.1) +
  geom_line(aes(colour = model_4), size = 1, color = "blue") +
  labs(x = "Day", y = "% of the Population Infected") +
  ggtitle("Average Cumulative Cases \nfor Median Household Income $46,500 - $54,000")


########## INCOME GROUP 5 ##########
exp_reg_5 <- lm(y_54001 ~ x_54001 + I(x_54001^2) + I(x_54001^3), data = xy_54001)

prd <- data.frame(x_54001 = seq(0, 350, by = 0.5))

result <- prd
result$exp_reg_5 <- predict(exp_reg_5, newdata = prd)

result <- melt(result, id.vars = "x_54001", variable.name = "model_5",
                value.name = "cases_5")
ggplot(result, aes(x = x_54001, y = cases_5)) +
  theme_bw() +
  geom_point(data = xy_54001, aes(x = x_54001, y = y_54001), size = 0.1) +
  geom_line(aes(colour = model_5), size = 1, color = "purple") +
  labs(x = "Day", y = "% of the Population Infected") +
  ggtitle("Average Cumulative Cases \nfor Median Household Income >$54,000")


########## Functions of the Regression Lines ##########
exp_reg_1 #Average Cumulative Cases for Median Household Income <$36,000
exp_reg_2 #Average Cumulative Cases for Median Household Income $36,000 - $41,000
exp_reg_3 #Average Cumulative Cases for Median Household Income $41,000 - $46,500
exp_reg_4 #Average Cumulative Cases for Median Household Income $46,500 - $54,000
exp_reg_5 #Average Cumulative Cases for Median Household Income >$54,000
```

Synthesizes and transforms the data so that we have a dataset that is best formatted for graphing the information we want to (date, 4-day case totals per capita, and income bracket), then graphs it

```{r, message=FALSE}
us_cases_per_capita <- data.frame(t(us_cases[-1]))
colnames(us_cases_per_capita) <- us_cases[, 1]
us_cases_per_capita <- us_cases_per_capita[, names(us_cases_per_capita) %in% income$County]

us_pop_filtered <- data.frame(t(us_pop[-1]))
colnames(us_pop_filtered) <- us_pop[, 1]
us_pop_filtered <- us_pop_filtered[, names(us_pop_filtered) %in% income$County]

b <- list(rep(0, 79), rep(0, 79), rep(0, 79), rep(0, 79), rep(0, 79))
l <- list(length(buckets[[1]]), length(buckets[[2]]), length(buckets[[3]]), length(buckets[[4]]), length(buckets[[5]]))

for(county in names(us_cases_per_capita))
  for(i in 1:5)
    if(county %in% buckets[[i]])
      b[[i]] <- b[[i]] +  colSums(matrix(c(us_cases_per_capita[[county]][1], diff(us_cases_per_capita[[county]])) / us_pop_filtered[[county]], nrow = 5))

dates <- rep(1:79, 5)
all_cases <- c(b[[1]] / l[[1]], b[[2]] / l[[2]], b[[3]] / l[[3]], b[[4]] / l[[4]], b[[5]] / l[[5]])
incomes <- c(rep(36000, 79), rep(41000, 79), rep(46500, 79), rep(54000, 79), rep(54001, 79))

bucketed_data <- data.frame(dates, all_cases, incomes)

p2 <- ggplot(bucketed_data, aes(dates, all_cases, group = incomes, color = factor(incomes))) +
  geom_line() +
  scale_color_viridis_d() +
  labs(x = "Month", y = "% of the Population Infected", color = "Median Household Income") +
  theme(legend.position = "top") + 
  scale_x_continuous(breaks = seq(7, 300, 29.3), labels=c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov")) +
  scale_color_discrete(breaks = c(36000, 41000, 46500, 54000, 54001), labels=c("< $36,000", "< $41,000", "< $46,500", "< $54,000", "> $54,000")) +
  ggtitle("Average Cases Every 4 Days") + 
  geom_point(aes(group = seq_along(dates))) + 
  transition_reveal(dates)

animate(p2, duration = 10, end_pause = 75)
```